---
title: "r4ds_chp21_markdown.Rmd"
author: "Shade Wilson"
date: "7/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

# Notes from R FOR DATA SCIENCE
an O'Reilly guide by Hadley Wickham and Garrett Grolemund. Availible online at http://r4ds.had.co.nz/

PART FIVE: Communicate

Chapter 21: R Markdown

cheat sheet availible at Help -> Cheatsheets -> R Markdown Cheat Sheet 

Cntrl/Command+Shift+M: `%>%` 

```{r iris_structure}
str(iris)
```

## Producing Reports
TO produce a complete report containing all text code and results, click "knit" or
press Cmd/Ctrl-Shift-K. Can also use `rmarkdown::render("1-example.Rmd")`

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Text Formatting with Markdown
*italic* or _italic_

**bold** or __bold__

`code`

superscript^2^ and subscript~2~

### Headings

# 1st level Header

## 2nd level Header

### 3rd level header

### Lists

* bulleted list item 1

* item 2

    * item 2a
  
    * item 2b
  
1. Numbered list item 1

1. Item 2. Numbers are incremented automatically in output

### Links and images

<https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf>

[R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)


### Tables

First Header | Second Header
------------ | -------------
Content Cell1| Content Cell2
Content cEll3| Content Cell4

## Code Chunks

Three ways to insert a chunk of code:

1. The keyboard chort cut Cmd/Ctrl-Alt-I

1. The "insert" button icon in the editor toolbar

1. By manually typing the chunk delimitors ```{r} and ```

YOu can run code using Cmd/Cntrl-Enter. Use Cmd/Ctrl+Shift+Enter to run all code in a chunk

### Chunk Name
```{r chunk_name}
name <- "chunk_name"
statement <- stringr::str_c("My name is ", name, ".")
print(statement)
```

Easier navigation, graphics produced by chunks will have more useful names, can set up
networks of cached chunks to avoid reperforming expensive computations on every run

chunk name "setup" has special behavior. When in notebook mode, the chunk named setup will be
run automatically once, before any other code is run

### Chunk options

Chunk options can be customized with *options*, args supplied to the chunk header. 
**knitr** provides almost 60 options to customize chunks

See full list at <http://yihui.name/knitr/options>

Most important set of options controls if your code block is executed adn what results 
are inserted in the finished report:

* `eval = FALSE` prevents code from being evaluated, therefore no results generated. Useful
for displaying example code or disabling a large block of code without using commenting

* `include = FALSE` runs the code but doesnt show the code or results in final document. Use
for setup code you dont want clittering your report

* `echo = FALSE` prevents code but not the results from appearing in finished file

* `message = FALSE` of `warning = FALSE` prevents messages or warnings to appear in finished
file

* `results = 'hide'` hides printed output; `fig.show = 'hide'` hides plots

* `error = TRUE` causes the render to continue even if code returns an error. Rare to use
in a final report but useful to debug whats going on inside your .Rmd. DEfault error = FALSE 
causes knit to fail if a single error exists

Following table summarizes which types of output each option suppresses:

Option | Run code | Show code | Output | Plots | Messages | Warnings
------ | -------- | --------- | ------ | ----- | -------- | --------
`eval = FALSE` | X |  | X | X | X | X
`include = FALSE` |  | X | X | X | X | X
`echo = FALSE` | | X | | | |
`results = 'hide'` | | | X | | |
`fig.show = 'hide'` | |  | | X | |
`message = FALSE` | | | | | X |
`warning = FALSE` | |  | | | | X

### Table

By defauly, R Markdown prints data frames and matrices as you'd see them in the consoles

```{r}
mtcars[1:5, 1:10]
```

If you'd prefer that data be displayed with additional fomratting, can use the `knitr::kable`
function

```{r knitr_kable}
knitr::kable(
  mtcars[1:5, ],
  caption = "A knitr kable"
)
```

Highly customizable. see `?knitr::kable`

### Caching

Normally each knit starts from a clean slate, which is good for reproducibility, but bad
if you have some computations that take a long time. Solution is `cache = TRUE`. When set, 
the output of chunk will be saved to a specially named file on disk. On subsequent runs,
knitr will check if the code has changed; if not, it will reuse cached results

Use with care! Based on code only, not dependencies

Can avoid this problem with the dependson chunk option: `dependson = "var_it_dependson"`

`dependson` should contain a character vector of every chunk that the catched chunk depends
on. knitr will update the results for the cached chunk whenever it detects that one of its
dependencies has changed.

Can clear all caches with `knitr::clean_cach()`

### Global Options

Can set default chunk options with `knitr::opts_chunk$set()` in a code chunk

Ex: when writing books and tutorials, Wickham sets:

```{r book_opts, eval = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
)

```

If preparing a report, you might set:
```{r report_opts, eval = FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)
```

This hides the code by default, only showing chunks that you deliberately want to show.

### Inline Code

One other way to embed R code into an R Markdown doc: directly into text with `r `

```{r smaller, include = FALSE}
library(ggplot2)
n <- 2.5
smaller <- dplyr::filter(diamonds, carat <= n)
```


Ex: We have data about `r nrow(diamonds)` diamonds. Only `r nrow(diamonds) - nrow(smaller)`
are larger than `r n` carats.

When inserting numbers into text, `format()` is your friend. Allows you to set the number of
digits so you dont print to a rediculous degree of accuracy. `big.mark` makes numbers
easier to read. Combined into a helper function:
```{r comma}
comma <- function(x) format(x, digits = 2, big.mark = ",")
comma(3442345)
comma(.123498627)
```


### Exercises

4.  
```{r a, cache = TRUE}
a <- runif(1)
print(lubridate::now())
```

```{r b, cache = TRUE}
b <- a + runif(1)
print(lubridate::now())
```

```{r c, cache = TRUE}
c <- a + 5
print(lubridate::now())
```

```{r d, cache = TRUE}
d <- c - b
print(lubridate::now())
```

## YAML Header

Can control many other "whole document" settings by tweaking the parameters of the YAML 
(yet another markup language) header.

### Parameters

Parameters are useful when you want to re-render the same report with distinct values for
various key inputs. TO declare one or more params, use `params` field.

EX:

output: html_document
params:
  my_class: "suv"
  
Can then use params$my_class to call specific class.

You can write atomic vectors directly into the YAML header. You can also run arbitrary
R expressions by prefacing the parameter value with !r. Good way to specify date/time params

params:
  start: !r lubridate::ymd("2015-01-01")
  snapshot: !r lubridate::ymd_hms("2015-01-01 12:30:00")
  
Alternatively, if you need to produce many such parameterized reports, can use
rmarkdown::render() with a list of params:

`rmarkdown::render("fuel-econmy.Rmd", params = list(my_class = "suv")`

Really good in conjunction with purrr:pwalk():

```{r pwalk_ex}
#rmarkdown::render("fuel-economy.Rmd", params = list(my_class = "suv"))

reports <- dplyr::tibble(
  class = unique(mpg$class),
  filename = stringr::str_c("fuel-economy-", class, ".html"),
  params = purrr::map(class, ~ list(my_class = .))
)
# reports %>%
#   select(output_file = filename, params) %>%
#   purrr::pwalk(rmarkdown::render, input = "fuel-economy.Rmd")
  
```

### Learn about Git and GitHub

<http://happygitwithr.com>

<http://r-pkgs.had.co.nz/git.html>