---
title: "r4ds_chp23_markdown_formats"
author: "Shade Wilson"
date: "8/4/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown Formats
Notes from R FOR DATA SCIENCE, an O'Reilly guide by Hadley Wickham and Garrett Grolemund. Availible online at <http://r4ds.had.co.nz/>. 

PART FIVE: Communicate

Chapter 23: R Markdown Formats

-----------------------------------------------------------------------------

Two ways to set the output of a document:

1. Permanently, by modifying the YAML header:

`title: "Viridis Demo"`

`output: html_document`

2. Transiently, by calling `rmarkdown::render()` by hand:
```{r, eval = FALSE}
rmarkdown::render(
  "diamond-sizes.Rmd",
  output_format = "word_document")
)
```

This is useful if you want to programmatically produce multiple types of output

### Output Options 
Each output format is associated with an R function. Can either write `foo` or `pkg::foo`. If you
omit `pkg`, the default is assumed to be **rmarkdown**.

Ex: `rmarkdown::html_document()`

TO override default parameters, need to use expanded output field. For example, if you wanted
to render an html_document with a floating table of contents, you'd use:

```{r}
# output:
#   html_document:
#     toc: true
#     toc_float: true
```

Can even render to multiple outputs by supplying a list of formats:
```{r}
# output:
#   html_document:
#     toc: true
#     toc_float: true
#   pdf_document: default
```

### Documents
Number of basic variations on documents (other than html).

* pdf_document makes a PDF with LaTeX (need to install).

* word_document for Word docs (.docx)

* odt_document for OpenDocument Text documents (.odt)

* rtf_document for RIch Text Format documents (.rtf)

* md_document for a Markdown document

* github_document is a tailored version of md_document for GitHub sharing

Remember, can turn off default display of code by setting global options in setup chunk: `knitr::opts_chunk$set(echo = FALSE)`

For html_documents, another option is to make the code chunks hudden by defaul, but visible with a
click:
```{r}
# output
#   html_document
#     code_folding: hide
```

### Notebooks
A notebook (html_notebook) is a variation on an html_document

blah

### Presentations
Can use R markdown to produce presentations.

### Dashboards
Useful way to communicate large amounts of information visually and quickly. **flexdashboard** makes
it particularly easy to create dashboards using R markdown. 

* each level 1 header (#) begins a new page

* each lvl 2 header begins a new column

* each lvl 3 header ebings a new row

### Shiny
add `runtime: shiny` to the header


